---
title: "Data Mining on Bank Marketing Data"
author: "Stephan Sprenger"
date: "25 October 2014"
output: html_document
---

# Introduction

The data investigated in this small workbook goes back to _S. Moro, P. Cortez and P. Rita. A Data-Driven Approach to Predict the Success of Bank Telemarketing. Decision Support Systems, Elsevier, 62:22-31, June 2014_  
The data is available from the [UCI Machine Learning Repository](https://archive.ics.uci.edu/ml/datasets/Bank+Marketing)

The purpose of this workbook to apply different machine learning algorithms to the data problem. The software ```R``` will be used in conjunction with the package ```caret```.


# Data Preparation

The data is being loaded. Investigation shows that the financial indicators are factor, while we would want to have them as numeric for analysis. Therefore this data is transformed.
We also understand that there are "unknown" data in several variables. We could replace these with NAs; however some algorithms (like Naive Bayes) would struggle. Given that on some of these variables there are many unknowns (like for ```pdays```) we will leave the unknown observations as an own category "unknown". 

The variable ```pdays``` has a value of 999 for 'not available'; wee will set this to "unknown" as well.

Also, the variable ```duration``` indicates the length of a call. this is interesting for analysis, but difficult for prediction: In a prediction task we do not know before the call how long it will take.  
Therefore ```duration``` is being excluded from the set of predictors; this seems to be different from a lot of other research on this data set.  

Also the data set is split into a training set (80%) and a test set (20%). 

```{r load data}
library(caret)
bank.raw <- read.csv2(file="bank-additional-full.csv")
str(bank.raw)
bank <- bank.raw
bank$job <- gsub("\\-", "", bank$job)
bank[,16:20] <- apply(bank[,16:20], 2, function (x) as.numeric(as.character(x)))
#bank[bank=="unknown"] <- NA
bank <- droplevels(bank)
bank[bank$pdays==999,"pdays"] <- "unknown"
bank$pdays <- as.factor(bank$pdays)
summary(bank)
names(bank)[names(bank)=="y"] <- "success"
bank$success <- factor(bank$success, levels=c("yes", "no"))

bank <- subset(bank, select=-c(duration))

trainIndex <- createDataPartition(bank$success, p=0.8, list=FALSE, times=1)
bank.train <- bank[trainIndex,]
bank.test <- bank[-trainIndex,]

```

# Data Exploration

Visual exploration 

```{r explo}
library(vcd)
mosaic(  ~  success+poutcome, data=bank, shade=TRUE, main="Success of campaign vs previous successes")
hist(bank$euribor3m)
library(ggplot2)
ggplot(data=bank, aes(x=euribor3m, fill=success)) + geom_histogram(alpha=0.5, position="dodge")
```

# Modelling

```{r modeling}
Sys.time()
bank.train.data <- subset(bank.train, select=-c(success))
bank.train.class <- subset(bank.train, select=c(success), drop=TRUE)

cvCtrl <- trainControl(method="cv", classProbs=TRUE, summaryFunction=twoClassSummary, verboseIter=TRUE)
#splCtrl <- trainControl(method="none", classProbs = TRUE)


fit.rpart <- train(bank.train.data, bank.train.class, method="rpart", trControl=cvCtrl, metric="ROC", tuneLength = 10)
#fit.rpart <- train(success ~ duration, data=bank, method="rpart")
fit.rpart
fit.rpart$finalModel
#summary(fit.rpart)
predict.rpart <- predict(fit.rpart)
confusionMatrix(predict.rpart, bank.train.class, positive="yes")
predict.rpart <- predict(fit.rpart, type = "prob")
(auc.rpart <- auc(bank.train.class, predict.rpart$yes))

fit.c50 <- train(bank.train.data, bank.train.class, method="C5.0", trControl=cvCtrl, metric="ROC", tuneLength = 10)
fit.c50
#fit.c50$finalModel
summary(fit.c50)
predict.c50 <- predict(fit.c50)
confusionMatrix(predict.c50, bank.train.class, positive="yes")
predict.c50 <- predict(fit.c50, type = "prob")
(auc.c50 <- auc(bank.train.class, predict.c50$yes))


fit.ctree <- train(bank.train.data, bank.train.class, method="ctree", trControl=cvCtrl, metric="ROC")
predict.ctree <- predict(fit.ctree)
confusionMatrix(predict.ctree, bank.train.class, positive="yes")
predict.ctree <- predict(fit.ctree, type = "prob")
(auc.ctree <- auc(bank.train.class, predict.ctree$yes))


fit.nb <- train(bank.train.data, bank.train.class, method="nb", trControl=cvCtrl, metric="ROC")
predict.nb <- predict(fit.nb)
confusionMatrix(predict.nb, bank.train.class, positive="yes")
predict.nb <- predict(fit.nb, type = "prob")
(auc.nb <- auc(bank.train.class, predict.nb$yes))

Sys.time()
fit.logit <- train(success ~ . , data=bank.train, method="glm", family=binomial, trControl=cvCtrl, metric="ROC", tuneLength = 10)
predict.logit <- predict(fit.logit)
confusionMatrix(predict.logit, bank.train.class, positive="yes")
predict.logit <- predict(fit.logit, type = "prob")
(auc.logit <- auc(bank.train.class, predict.logit$yes))

x <- glm(success~., family=binomial, data=bank.train)
predict(x)

fit.lr <- train(bank.train.data, bank.train.class, method="mlp", trControl=cvCtrl, metric="ROC")
1predict.lr <- predict(fit.nb)
confusionMatrix(predict.nb, bank.train.class, positive="yes")
predict.nb <- predict(fit.nb, type = "prob")
(auc.nb <- auc(bank.train.class, predict.nb$yes))



plot.roc(bank.train.class, predict.rpart$yes, col="green")
lines.roc(bank.train.class, predict.c50$yes, col="blue")
lines.roc(bank.train.class, predict.ctree$yes, col="pink")
lines.roc(bank.train.class, predict.nb$yes, col="red")
lines.roc(bank.train.class, predict.logit$yes, col="orange")
legend("bottomright", legend=c("CART", "C50", "Cond. Inference Trees", "Naive Bayes", "Logit"), col=c("green", "blue", "pink", "red", "orange"), lwd=2)

Sys.time()
plot(varImp(fit.rpart))
plot(varImp(fit.c50))
plot(varImp(fit.ctree))

```



```{r compare}
#compare different models
modelObjects <- list(cart=fit.rpart,
                     c50=fit.c50,
                     ctree=fit.ctree)
#predict(modelObjects, churnTrain)

x <- data.frame(ROC=sort(sapply(modelObjects, function(x) max(x$results$ROC)), decreasing=TRUE))
x
# variances in resamples.
cvValues <- resamples(list(CART=fit.rpart, C50=fit.c50, ctree=fit.ctree))
summary(cvValues)
dotplot(cvValues)
```

